Geometric variational problems

### 2020-11-06 NSF2020/Description/specificaim3_ver2.tex
small edits

\subsection{Specific Aim 3: Efficient high-order numerical methods for
large-scale simulations}
\label{subsec:specific_aim_3}
% ----------------------------------------------------------------------
The HAP model requires solving the exterior Dirichlet problem of the
screened Laplace equation and the mobility problem of the Stokes flow at
each time step in the complex domains such as Figure~\ref{fig:domain}.
If discretizing these equations with stencil-based numerical methods,
the computational domain must be truncated, the volume must be
discretized, and artificial boundary conditions must be imposed. The
artificial boundary conditions introduces additional error, and
discretizing the volume results in an excessively large linear system.
Moreover, it is difficult to obtain high-order discretizations when the
boundary surfaces are irregular. Since we are solving
constant-coefficient elliptic PDEs, we represent the solution of the PDE
of interest as a layer potential 
\begin{align}
  \label{eq:LP}
  u(x) = \int_{\partial\Omega} K(x,y) \sigma(y) ds_y,
\end{align}
where $\sigma$ is an unknown density function, and $K(x,y)$ is formed
with derivatives of the fundamental solution of the underlying PDE. By
matching the boundary condition at $x_0 \in \partial\Omega$ with the
limit of equation~\eqref{eq:LP} as $x\rightarrow x_0 \in
\partial\Omega$, a BIE for the density function is formed.
Numerical methods to solve BIE has several advantages over their
PDE-based counterpart: by construction, the layer
potential~\eqref{eq:LP} satisfies far-field conditions; only
$\partial\Omega$ needs to be discretized; carefully chosen kernels
$K(x,y)$ result in a well-conditioned linear system that requires a
mesh-independent number of GMRES iterations to solve; and high-order or
spectral accuracy is attainable by using appropriate quadrature methods.
In our previous work~\cite{Fu2018_SIAM} we demonstrate that BIEs are a
powerful tool to simulate two-dimensional suspensions of amphiphilic
particles. In addition to improving two-dimensional simulations, we will
extend the results to three dimensions using a standard
three-dimensional BIE of the screened Laplace equation~\cite{ying_2006}
and a well-conditioned BIE of the mobility problem~\cite{manasthesis}.

>>>>

\subsection{Specific Aim 3: Efficient high-order numerical methods for
large-scale simulations}
\label{subsec:specific_aim_3}
% ----------------------------------------------------------------------
The HAP model requires solving exterior Dirichlet problems for the
screened Laplace equation and the Stokes mobility problem at each time
step in complex domains such as in Figure~\ref{fig:domain}. If
discretizing these equations with stencil-based numerical methods, the
computational domain must be truncated, the fluid volume must be
discretized, and artificial boundary conditions must be imposed. The
artificial boundary conditions introduces additional error, and
discretizing the volume results in an excessively large linear system.
Moreover, it is difficult to obtain high-order discretizations when the
boundary surfaces are irregular. Alternatively, since we are solving
constant-coefficient elliptic PDEs, the solution of the PDE can be
represented as layer potentials
\begin{align}
  \label{eq:LP}
  u(x) = \int_{\partial\Omega} K(x,y) \sigma(y)\,ds_y,
\end{align}
where $\sigma$ is an unknown density function, and $K(x,y)$ is a
derivative of the fundamental solution of the underlying PDE. By
matching the boundary condition at $x_0 \in \partial\Omega$ with the
limit of equation~\eqref{eq:LP} as $x\rightarrow x_0$, a BIE for the
density function is formed. Numerical methods to solve BIEs have several
advantages over their PDE-based counterpart: by construction, the layer
potential~\eqref{eq:LP} satisfies far-field conditions; only
$\partial\Omega$ needs to be discretized; carefully chosen kernels
$K(x,y)$ result in a well-conditioned linear system that is solved with
a mesh-independent number of GMRES iterations; and high-order or
spectral accuracy is attainable with appropriate quadrature methods. In
our previous work~\cite{Fu2018_SIAM} we demonstrate that BIEs are a
powerful tool to simulate two-dimensional suspensions of amphiphilic
particles. In addition to improving two-dimensional simulations, we will
extend the results to three dimensions using a standard
three-dimensional BIE of the screened Laplace equation~\cite{ying_2006}
and a well-conditioned BIE of the mobility problem~\cite{manasthesis,
rac-gre2016}.
